Bayesian exponential random graph modeling of KWL Burial network

### 2020-11-11 analysis/paper/paper.Rmd
add references and citations

<!-- two types of network analysis -->
Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017]. Another approach is using models and simulations to represent and test dependence of relationships [@Brughmans2018]. Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns. Recent applications include the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Similarly, @Mizoguchi2013 analysed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods. For the second approach, researchers typically statistically test hypotheses about the relationships between actors to examine the structure of the network and infer major factors that affect the formation of the structure [@Freeman2004; @Salvini2010]. Examples include testing hunter-gatherer exchange networks across the Kuril Islands based on ceramic composition [@Erik2015].  <!-- better to have a specific case study by a scholar who is likely to peer review this paper, try to find some recent publication --> . This paper will use a novel Bayesian modeling approach to investigate the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organisation.

<!-- archaeological background -->
Archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007]. This might indicate the pursuit of prestige or wealth in those Indigenous societies. Prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks that involved the Dutch, Spanish, Han Chinese traders, and local Indigenous groups. The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period <!-- citation -->. For example, an Indigenous person possessing more imported goods may have been recognised as having higher status by their community <!-- citation to some historical record -->. A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks <!-- citation -->.

>>>>

<!-- two types of network analysis -->
Network analysis has been increasingly applied by archaeologists in recent years to deal with past interactions and explore the underlying mechanisms. There are two common approaches, one is directly examining network properties, such as node position and overall structures of networks built on archaeological data, to find meaningful patterns [@Mills2017]. Another approach is using models and simulations to represent and test dependence of relationships [@Brughmans2018]. Node position is important to identify the role of a node in a network, such as its centrality, representing the individual influence or social prominence in a group, while network structure is helpful to generalize relationship patterns. Recent applications include the investigation of long term inter-site relationships from the Epipalaeolithic to the early Neolithic in the Near East according to trade items [@Coward2013]. Similarly, @Mizoguchi2013 analyzed hierarchical communication systems in the Kofun period in Japan through the study of prestige goods. Regarding burials, @Sosna examined spatial pattern of burials from the Early Bronze Age in Rebesovice with two hypothesized networks constructed according to cultural and chronological differences between burials. They found that the observed concentric pattern does not match the network results projected into geographic space, but contemporary subsections that indicates different corporate groups. For the second approach, researchers typically statistically test hypotheses about the relationships between actors to examine the structure of the network and infer major factors that affect the formation of the structure [@Freeman2004; @Salvini2010]. Examples include accessing hunter-gatherer exchange networks structure across the Kuril Islands using bootstrap simulation based on ceramic composition [@Erik2015]. Another example is to simulate several modes of food exchange for Ancestral Pueblos and compare the results with the social networks build on the aggregation of households in the American Southwest at a regional scale [@Crabtree2015]. This paper will use a novel Bayesian modeling approach to investigate the relationship between burials in northeastern Taiwan to explore the indirect impact of foreign contacts on social organisation.

<!-- archaeological background -->
Archaeological sites in northeastern Taiwan show evidence of use of imported prestige goods in burials and some accumulation patterns in residential areas around the same time as European contact in the 17th century [@Chen2007]. This might indicate the pursuit of prestige or wealth in those Indigenous societies. Prestige goods include imported ornaments and porcelains that entered northeastern Taiwan via regional exchange networks that involved the Dutch, Spanish, Han Chinese traders, and local Indigenous groups. The use of prestige goods may have began in local Indigenous communities before a European presence became established, but was amplified during the European colonial period [@Borao2009; @LiandWu2006; @LWandBM2020ornament]. For example, an Indigenous person possessing more imported goods may have been recognized as more influential or having higher status by their community [@LiandWu2006]. A general model to summarize this situation in northeastern Taiwan might be that the influence of a colonial power, combined with high local values of imported goods, might lead to increased social inequality due to competition among individuals, and unequal access to trade networks <!-- citation -->.
